---
title: "p"
author: "me"
date: "3/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
install.packages("naniar")
install.packages("skimr")
```
ADD README and DESCRIPTION
FINISH BY FRIDAY
awsome readme
```{r}
library(tidyverse)
library(shiny)
library(shinydashboard)
```
```{r}
AbaloneData<-readr::read_csv("AbaloneData.csv")
AbaloneData
##I renamed the data to AbaloneData and moved it into a folder with this r code and stored it in a value also named AbaloneData
``` 
```{r}
naniar::miss_var_summary(AbaloneData)
```
make generic app that explores possible correlations between all factors and maximum growth
      i.e. x factor by growth
extra: use calculus functions to calculate the optimal tradeoff between growth and some other factor like spawn rate or costs of raising
potential presentation format
  1. introduce abalone/why are abalone important to us(food,ecosystem,etc)
  2. explain collected data
  3. show analysis (apps)
    plots/apps to show correlations with mass and through
    correlation between sex and mass
    plots/apps to show correlation to sex
    show fertility(gonad score > 0)
      (gonad score vs mass & length)
      fertility vs sex
      length vs mass filled by predicted sex
      average growth -> use to compare to other factors like sex and gonad score and trough
  4. conclusion/further projects-ecosystem,growthrates, does mass focus affect sex