---
title: "p"
author: "me"
date: "3/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
install.packages("naniar")
install.packages("skimr")
```
```{r}
library(tidyverse)
library(shiny)
library(shinydashboard)
```
```{r}
AbaloneData<-readr::read_csv("AbaloneData.csv")
AbaloneData
##I renamed the data to AbaloneData and moved it into a folder with this r code and stored it in a value also named AbaloneData
``` 
```{r}
glimpse(AbaloneData)
naniar::miss_var_summary(AbaloneData)
```
make generic app that explores possible correlations between all factors and maximum growth
      I.e. x factor by growth
extra: use calculus functions to calculate the optimal tradeoff between growth and some other factor like spawn rate or costs of raising
potential presentation format
  1. introduce abalone
  2. why are abalone important to us(food,ecosystem,etc)
  3. explain collected data
  4. show analysis (apps)
  5. conclusion